---
title: "Multivariate Descriptive Statistics"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
```

```{r load, message=FALSE, warning=FALSE}
# Loading the dataset into a dataframe
df <- read_delim("../../data/processed/wines.csv", 
  ";", 
  escape_double = FALSE, 
  trim_ws = TRUE)
```

# Analysis of the dataset
## Quantitative attributes

```{r echo=FALSE}
summary(df)
```

\begin{table}[]
\centering
\caption{Analysis of the attributes.}
\label{my-label}
\begin{tabular}{|l|c|c|c|c|}
\hline
                       & \textbf{min} & \textbf{max} & \textbf{mean} & \textbf{SD} \\ \hline
fixed\_acidity         & 3,800        & 15,900       & 7,215         & 1.296588    \\ \hline
volatile\_acidity      & 0,0800       & 1,5800       & 0,3396        & 0.164583    \\ \hline
citric\_acid           & 0,0          & 1,6600       & 0,3187        & 0.1452326   \\ \hline
residual\_sugar        & 0,600        & 65,800       & 5,444         & 4.758494    \\ \hline
chlorides              & 0,00900      & 0,61100      & 0,05602       & 0.03503299  \\ \hline
free\_sulfur\_dioxide  & 1,00         & 289,00       & 30,52         & 17.74849    \\ \hline
total\_sulfur\_dioxide & 6,0          & 440,0        & 115,8         & 56.52657    \\ \hline
density                & 0,9871       & 1,0390       & 0,9947        & 0.002999095 \\ \hline
pH                     & 2,720        & 4,010        & 3,219         & 0.1608116   \\ \hline
sulphates              & 0,2200       & 2,00         & 0,5313        & 0.148822    \\ \hline
alcohol                & 8,00         & 14,90        & 10,49         & 1.192768    \\ \hline
\end{tabular}
\end{table}


## Quality histogram

```{r echo=FALSE}
type.0 = subset(df, type==0)
type.1 = subset(df, type==1)
type.0$type <- 'white'
type.1$type <- 'red'
types <- rbind(type.0, type.1)
ggplot(types, aes(types$quality, fill = type)) + geom_histogram(alpha = 0.5)+ ggtitle('Histogram of wines quality: red vs. white')+ylab("Wine samples")+xlab("Quality")
#hist(type.0$quality, main="White wine vs. Red wine quality", xlab="Quality", ylab="Wine samples")
#hist(type.1$quality, col="red")
#box()

```

## Correlation analysis
Análisis de `correlations` (function corrplot.mixed with first argument r) y `partial correlations` (function corrplot.mixed with first argument partial.corr).
```{r echo=FALSE}

```

## Coefficients of determination
Para las variables numéricas (no categóricas):s Interesante también calcular los `coefficients of determination` de cada variable (function r2multv) y la `effective dependence coefficient of the R matrix` (function 1-det(cor(ds[,c(1,5,7,8,9,10)]))^(1/6)) El 6 de 1/6 es el número de atributos. 
```{r echo=FALSE}

```



